---
title: "&#32;"
output:
  html_document:
    toc: false
    toc_float: false
    number_sections: false
    df_print: paged
    self_contained: true
    theme: united

params:
  var1: null
  var2: null
  var1_path: null
  var2_path: null
  Methylome.Plants_path: "."

  annotation_file: null
  description_file: null
  TEs_file: null
  reference_bundle_path: null
  species_name: "Plant"
  assembly_name: "custom"

  minProportionDiff: [0.4, 0.2, 0.1]    # CG, CHG, CHH
  binSize: 100
  minCytosinesCount: 4
  minReadsPerCytosine: 4
  pValueThreshold: 0.05
  methyl_files_type: "CX_report"
  img_type: "png"
  n.cores: 8

  analyze_DMRs: true
  run_QC: true
  run_PCA_plot: true
  run_total_meth_plot: true
  run_CX_Chrplot: true
  run_TEs_distance_n_size: true
  total_meth_annotation: true
  run_TF_motifs: true
  run_functional_groups: true
  run_GO_analysis: false
  run_KEGG_pathways: false

  analyze_strand_asymmetry_DMRs: false
  analyze_DMVs: false
  analyze_dH: false
  run_TE_metaPlots: false
  run_GeneBody_metaPlots: false
  run_GeneFeatures_metaPlots: false
  gene_features_binSize: 10
  metaPlot.random.genes: 10000

  # ---------------------------
  # Report layout (images only)
  # ---------------------------
  plot_max_width_px: 1100      # max width for each plot
  plot_max_height_px: 500      # cap plot height (prevents huge SVG/PNG)
  plot_gap_px: 5              # gap between plots in grids
  default_plot_cols: 1         # default columns for multi-plot sections

  # If TRUE: copy plots next to the HTML (robust links even when HTML is saved inside results/<comparison>)
  copy_images_next_to_report: true
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressPackageStartupMessages({ library(knitr) })

# Wrapper: run code block, catch any error and print a warning instead of stopping
safe_section <- function(expr) {
  tryCatch(expr, error = function(e) {
    cat(sprintf("\n\n<div style='color:#cb0a0a; border:1px solid #cb0a0a; padding:8px; border-radius:4px;'>**Error rendering section:** %s</div>\n\n", conditionMessage(e)))
  })
}

stop_if_missing <- function(x, name) {
  if (is.null(x) || length(x) == 0 || !nzchar(as.character(x)[1])) stop("Missing required param: ", name)
}
stop_if_missing(params$var1, "var1")
stop_if_missing(params$var2, "var2")

root <- params$Methylome.Plants_path
if (!dir.exists(root)) stop("Methylome.Plants_path does not exist: ", root)
knitr::opts_knit$set(root.dir = root)

comparison_name <- paste0(params$var2, "_vs_", params$var1)
exp_path <- file.path("results", comparison_name)
if (!dir.exists(file.path(root, exp_path)) && !dir.exists(exp_path)) {
  # When root.dir is set, relative paths are resolved from it. This check is defensive.
  stop("Results folder not found: ", exp_path)
}

# Output folders (match Methylome.Plants_main.R)
qc_dir_path <- file.path(exp_path, "QC")
total_meth_path <- file.path(exp_path, "total_methylation_analysis")
PCA_plots_path <- file.path(total_meth_path, "PCA_plots")
meth_levels_path <- file.path(total_meth_path, "methylation_levels")
ChrPlot_CX_path <- file.path(total_meth_path, "ChrPlot_CX")
ChrPlot_subCX_path <- file.path(ChrPlot_CX_path, "subCX")
TEs_distance_n_size_path <- file.path(total_meth_path, "TE_size_n_distance")
total_meth_annotation_path <- file.path(total_meth_path, "total_methylation_annotations")
TF_motifs_path <- file.path(total_meth_path, "TF_motifs")

DMRs_analysis_path <- file.path(exp_path, "DMR_analysis")
gainORloss_path <- file.path(DMRs_analysis_path, "gain_OR_loss")
genome_ann_path <- file.path(DMRs_analysis_path, "genome_annotation")
ChrPlots_DMRs_path <- file.path(DMRs_analysis_path, "ChrPlot_DMRs")
DMRs_bigWig_path <- file.path(DMRs_analysis_path, "DMRs_bigWig")

strand_asymmetry_path <- file.path(exp_path, "Strand_Asymmetry_DMRs")
DMV_analysis_path <- file.path(exp_path, "DMV_analysis")
dH_path <- file.path(exp_path, "deltaH")
dH_ann_path <- file.path(dH_path, "genome_annotation")
meta_dir <- file.path(exp_path, "MetaPlots")

# ----------------------------
# Plot rendering utilities (IMAGES ONLY: SVG/PNG/JPEG/TIFF)
# ----------------------------

# Determine where the HTML will be written; used for copying plot assets near the report.
get_report_dir <- function() {
  od <- tryCatch(knitr::opts_knit$get("output.dir"), error = function(e) NULL)
  if (!is.null(od) && nzchar(od)) return(normalizePath(od, winslash = "/", mustWork = FALSE))
  co <- tryCatch(knitr::current_output(), error = function(e) NULL)
  if (!is.null(co) && nzchar(co)) return(dirname(normalizePath(co, winslash = "/", mustWork = FALSE)))
  # common default when output_file is inside results/<comparison>
  normalizePath(exp_path, winslash = "/", mustWork = FALSE)
}
report_dir <- get_report_dir()

# Where we store image assets for stable linking
assets_dir <- file.path(report_dir, paste0(comparison_name, "_plots"))
if (isTRUE(params$copy_images_next_to_report)) dir.create(assets_dir, recursive = TRUE, showWarnings = FALSE)

# Convert an absolute or relative path to a relative path under project root (root.dir)
project_rel <- function(p) {
  if (is.null(p) || !nzchar(p)) return(p)
  # If already relative, keep it
  if (!grepl("^(/|[A-Za-z]:)", p)) return(p)
  wd <- tryCatch(normalizePath(getwd(), winslash = "/", mustWork = FALSE), error = function(e) "")
  pa <- tryCatch(normalizePath(p, winslash = "/", mustWork = FALSE), error = function(e) p)
  if (nzchar(wd) && startsWith(pa, paste0(wd, "/"))) sub(paste0("^", wd, "/"), "", pa) else pa
}

# Copy an image into assets_dir while preserving its relative path under exp_path
asset_link <- function(img_file) {
  if (!file.exists(img_file)) return(NA_character_)
  # ignore PDFs completely (report is non-PDF)
  if (tolower(tools::file_ext(img_file)) == "pdf") return(NA_character_)

  if (!isTRUE(params$copy_images_next_to_report)) return(chartr("\\", "/", project_rel(img_file)))

  exp_abs <- normalizePath(exp_path, winslash = "/", mustWork = FALSE)
  img_abs <- normalizePath(img_file, winslash = "/", mustWork = FALSE)

  rel <- if (startsWith(img_abs, paste0(exp_abs, "/"))) sub(paste0("^", exp_abs, "/"), "", img_abs) else basename(img_abs)
  dst <- file.path(assets_dir, rel)
  dir.create(dirname(dst), recursive = TRUE, showWarnings = FALSE)
  file.copy(img_abs, dst, overwrite = TRUE)

  # return link relative to report_dir
  rd_abs <- normalizePath(report_dir, winslash = "/", mustWork = FALSE)
  dst_abs <- normalizePath(dst, winslash = "/", mustWork = FALSE)
  link <- if (startsWith(dst_abs, paste0(rd_abs, "/"))) sub(paste0("^", rd_abs, "/"), "", dst_abs) else dst_abs
  chartr("\\", "/", link)
}

is_media <- function(x) {
  ext <- tolower(tools::file_ext(x))
  ext %in% c("png","jpg","jpeg","svg","tiff")
}

list_media_files <- function(dir_path, recursive = FALSE, pattern = NULL) {
  if (!dir.exists(dir_path)) return(character())
  files <- list.files(dir_path, full.names = TRUE, recursive = recursive)
  files <- files[!file.info(files)$isdir]
  files <- files[sapply(files, is_media)]
  if (!is.null(pattern)) files <- files[grepl(pattern, basename(files), ignore.case = TRUE)]
  files
}

context_rank <- function(x) {
  b <- tolower(basename(x))
  if (grepl("all_contexts|allcontexts|all", b)) return(0L)
  if (grepl("\\bcg\\b", b)) return(1L)
  if (grepl("\\bchg\\b", b)) return(2L)
  if (grepl("\\bchh\\b", b)) return(3L)
  9L
}

order_files <- function(files) {
  if (length(files) == 0) return(files)
  files[order(vapply(files, context_rank, integer(1)), tolower(basename(files)))]
}

emit_grid <- function(files, cols = params$default_plot_cols, max_height_px = NULL) {
  files <- files[file.exists(files)]
  files <- order_files(files)
  if (length(files) == 0) {
    cat("_No plot files found._

")
    return(invisible(NULL))
  }

  cols <- as.integer(cols)
  if (is.na(cols) || cols < 1L) cols <- 1L

  mh <- max_height_px
  if (!is.null(mh)) {
    mh <- as.integer(mh)
    if (is.na(mh) || mh < 50L) mh <- NULL
  }

  if (is.null(mh)) {
    cat(sprintf('<div class="plot-grid" style="--cols:%d;">
', cols))
  } else {
    cat(sprintf('<div class="plot-grid" style="--cols:%d; --img-max-height:%dpx;">
', cols, mh))
  }

  for (f in files) {
    src <- asset_link(f)
    if (is.na(src) || !nzchar(src)) next
    cat(sprintf('  <div class="plot-cell"><img src="%s" loading="lazy"></div>
', src))
  }
  cat('</div>

')
  invisible(NULL)
}

emit_dir <- function(dir_path, recursive = FALSE, cols = params$default_plot_cols, pattern = NULL, max_height_px = NULL) {
  files <- list_media_files(dir_path, recursive = recursive, pattern = pattern)
  emit_grid(files, cols = cols, max_height_px = max_height_px)
}

# Compact printing of params that may be vectors
val1 <- function(x) {
  if (is.null(x) || length(x) == 0) return(NA_character_)
  paste(as.character(x), collapse = ", ")
}
```

```{r plot-style, results='asis', echo=FALSE}
cat(sprintf('<style>
:root{
  --plot-max-width: %dpx;
  --plot-max-height: %dpx;
  --plot-gap: %dpx;
}
.plot-grid{
  display:flex;
  flex-wrap:wrap;
  gap: var(--plot-gap);
  justify-content: center;
  align-items:flex-start;
  margin: 10px 0 18px 0;
  --img-max-height: var(--plot-max-height);
}
.plot-cell{
  flex: 0 0 calc((100%% - (var(--cols) - 1) * var(--plot-gap)) / var(--cols));
  max-width: var(--plot-max-width);
}
.plot-cell img{
  width:100%%;
  height:auto;
  max-height: var(--img-max-height);
  object-fit: contain;
  display:block;
}
</style>', as.integer(params$plot_max_width_px), as.integer(params$plot_max_height_px), as.integer(params$plot_gap_px)))
```

[Methylome.Plants](https://github.com/Yo-yerush/Methylome.Plants) report. *`r format(Sys.Date(), '%d, %B, %Y')`*

## `r params$var2` *vs.* `r params$var1` results overview {.tabset}

### Introduction

This report summarizes the results of the **Methylome.Plants** pipeline for **`r params$species_name`** (assembly **`r params$assembly_name`**). The pipeline performs quality control, computes global and context-specific DNA methylation levels, identifies differentially methylated regions (DMRs), and generates visualizations and summary tables for reproducible methylome comparisons.

**Key features of the pipeline:**

- Automated QC and conversion rate estimation
- Global and per-context methylation profiling
- Chromosome-scale and annotation-based methylation summaries
- DMR detection and annotation
- Visualization of methylation patterns, PCA, and functional group overlaps
- Optional analyses: motif enrichment, GO/KEGG, meta-plots, and more

---

#### Analyses checklist

```{r analyses-checklist, results='asis', echo=FALSE}
check <- function(flag) if (isTRUE(flag)) "[x]" else "[ ]"

cat(sprintf("
**QC**

- %s Sample-level QC plots

**Total methylation**

- %s PCA (Principal Component Analysis)
- %s Total methylation levels
- %s Chromosome methylation profiles
- %s TE methylation vs TE size & centromere distance
- %s Total methylation across genome annotations
- %s TF motif analysis

**DMR analysis**

- %s DMR detection & annotation
- %s Functional groups
- %s GO analysis
- %s KEGG pathways

**MetaPlots**

- %s TE MetaPlots
- %s Gene body MetaPlots
- %s Gene features MetaPlots

**Other analyses**

- %s Strand-asymmetry DMRs
- %s DMVs (Differentially Methylated Valleys)
- %s ΔH / entropy (SurpDMRs)
",
check(params$run_QC),
check(params$run_PCA_plot),
check(params$run_total_meth_plot),
check(params$run_CX_Chrplot),
check(params$run_TEs_distance_n_size),
check(params$total_meth_annotation),
check(params$run_TF_motifs),
check(params$analyze_DMRs),
check(params$run_functional_groups),
check(params$run_GO_analysis),
check(params$run_KEGG_pathways),
check(params$run_TE_metaPlots),
check(params$run_GeneBody_metaPlots),
check(params$run_GeneFeatures_metaPlots),
check(params$analyze_strand_asymmetry_DMRs),
check(params$analyze_DMVs),
check(params$analyze_dH)
))
```

### QC {.tabset}

**Comparison:** `r params$var2` vs `r params$var1`  
**Results folder:** `r file.path(root, exp_path)`

#### Conversion rate

When a chloroplast sequence is configured in the reference bundle, its largely unmethylated DNA provides a practical estimate of bisulfite C→T conversion efficiency.

```{r conversion-rate, results='asis'}
safe_section({
  conv_path <- file.path(qc_dir_path, "conversion_rate.csv")
  if (!file.exists(conv_path)) {
    cat(sprintf("_Missing file:_ `%s`_\n\n", conv_path))
  } else {
    conv <- read.csv(conv_path, check.names = FALSE)
    space_rep <- rep(" ", nrow(conv))
    conv <- cbind(conv, data.frame(a=space_rep, b=space_rep, c=space_rep, d=space_rep, e=space_rep))
    colnames(conv) <- c("Sample", "Conversion (%)", "", "", "", "", "")
    kable(conv, align = c("l", "l"))
  }
})
```

```{r qc-plots-section, results='asis'}
if (isTRUE(params$run_QC) && dir.exists(qc_dir_path)) { safe_section({

  qc_dist_path <- file.path(qc_dir_path, "distribution")
  qc_scatter_path <- file.path(qc_dir_path, "scatter")
  qc_corr_path <- file.path(qc_dir_path, "correlation")
  qc_var_path <- file.path(qc_dir_path, "variance")

  cat("#### Methylation distributions\n\n")
  cat("Per-region CG/CHG/CHH methylation distributions for gene bodies, promoters and TEs (control vs treatment).\n\n")
  if (dir.exists(qc_dist_path)) {
    dist_files <- list_media_files(qc_dist_path, pattern = "QC_meth_distribution_")
    if (length(dist_files) > 0) {
      emit_grid(dist_files, cols = 1, max_height_px = params$plot_max_height_px * 1.5)
    } else {
      cat("_No distribution plots found._\n\n")
    }

    dist_csv <- list.files(qc_dist_path, pattern = "QC_meth_distribution_summary_.*\\.csv$", full.names = TRUE)
    if (length(dist_csv) > 0) {
      dist_summary <- read.csv(dist_csv[1], check.names = FALSE)
      cat("\n")
      print(kable(dist_summary))
      cat("\n\n")
    }
  }

  cat("#### Sample scatter plots {.tabset}\n\n")
  cat("Pairwise sample-vs-sample scatter (hexbin) plots on genomic tiles. Pearson r is shown per pair.\n\n")
  if (dir.exists(qc_scatter_path)) {
    scatter_files <- list_media_files(qc_scatter_path, pattern = "QC_scatter_")
    if (length(scatter_files) > 0) {
      for (cntx_s in c("CG", "CHG", "CHH")) {
        cntx_scatter <- scatter_files[grepl(paste0("QC_scatter_", cntx_s, "_"), basename(scatter_files))]
        if (length(cntx_scatter) > 0) {
          cat(paste0("##### ", cntx_s, "\n\n"))
          emit_grid(cntx_scatter, cols = 5, max_height_px = params$plot_max_height_px)
        }
      }
    } else {
      cat("_No scatter plots found._\n\n")
    }
  }

  cat("#### Sample correlation\n\n")
  cat("Pearson correlation heatmap across all replicates per context.\n\n")
  if (dir.exists(qc_corr_path)) {
    corr_files <- list_media_files(qc_corr_path, pattern = "QC_correlation_heatmap_")
    if (length(corr_files) > 0) {
      emit_grid(order_files(corr_files), cols = 3, max_height_px = params$plot_max_height_px)
    } else {
      cat("_No correlation heatmaps found._\n\n")
    }

    corr_csvs <- list.files(qc_corr_path, pattern = "QC_correlation_matrix_.*\\.csv$", full.names = TRUE)
    if (length(corr_csvs) > 0) {
      for (cc in corr_csvs) {
        cntx_label <- sub(".*QC_correlation_matrix_(CG|CHG|CHH)_.*", "\\1", basename(cc))
        corr_mat <- read.csv(cc, check.names = FALSE, row.names = 1)
        cat(paste0("\n**", cntx_label, " correlation matrix**\n\n"))
        print(kable(round(corr_mat, 4)))
        cat("\n\n")
      }
    }
  }

  cat("#### Variance / dispersion test\n\n")
  cat("Per-tile variance comparison between conditions. Tests whether treatment variance is systematically higher than control (paired Wilcoxon).\n\n")
  if (dir.exists(qc_var_path)) {
    var_files <- list_media_files(qc_var_path, pattern = "QC_variance_")
    if (length(var_files) > 0) {
      emit_grid(var_files, cols = 3, max_height_px = params$plot_max_height_px)
    } else {
      cat("_No variance plots found (requires >= 2 replicates per condition)._\n\n")
    }

    var_csv <- list.files(qc_var_path, pattern = "QC_variance_test_.*\\.csv$", full.names = TRUE)
    if (length(var_csv) > 0) {
      var_table <- read.csv(var_csv[1], check.names = FALSE)
      cat("\n")
      print(kable(var_table))
      cat("\n\n")
    }
  }
})}
```

```{r total-meth-tabs, results='asis'}
any_total_meth <- isTRUE(params$run_PCA_plot) || isTRUE(params$run_total_meth_plot) ||
  isTRUE(params$run_CX_Chrplot) || isTRUE(params$run_TEs_distance_n_size) ||
  isTRUE(params$total_meth_annotation) || isTRUE(params$run_TF_motifs)

if (any_total_meth) { safe_section({
  cat("### Total methylation {.tabset}\n\n")

if (isTRUE(params$run_PCA_plot)) {
  cat("#### PCA (Principal Component Analysis)\n\n")
  cat("PCA compresses genome-wide methylation into a few axes. Replicates should cluster; separation between conditions suggests global methylation shifts.\n\n")
  emit_dir(PCA_plots_path, recursive = FALSE, cols = 4, max_height_px = params$plot_max_height_px)
}

if (isTRUE(params$run_total_meth_plot)) {
  cat("#### Total methylation levels\n\n")
  cat("Global methylation (%) summarized per context and condition.\n\n")
  emit_dir(meth_levels_path, recursive = FALSE, cols = 3, max_height_px = params$plot_max_height_px/1.5)
}

if (isTRUE(params$run_CX_Chrplot)) {
  cat("#### Chromosome methylation profiles {.tabset}\n\n")
  cat("Chromosome-scale methylation. Often highlights pericentromeric/heterochromatin-linked changes.\n\n")

  ## Split into context subtabs if possible
  all_files <- list_media_files(ChrPlot_CX_path, recursive = FALSE)
  if (length(all_files) == 0) {
    cat("_No chromosome methylation plots found._\n\n")
  } else {
  
      for (f in all_files) {
        chrplot_filename <- gsub(paste0(params$var2, "_vs_", params$var1), "", tools::file_path_sans_ext(basename(f)))
        chrplot_filename <- gsub("_", " ", chrplot_filename)
        cat("##### ", chrplot_filename, "\n\n", sep="")
        emit_grid(f, cols = 1, max_height_px = params$plot_max_height_px)
      }
  }

  if (dir.exists(ChrPlot_subCX_path)) {
    sub_files <- list_media_files(ChrPlot_subCX_path, recursive = TRUE)
    cat("##### subCX {.tabset}\n\n")
    if (length(sub_files) > 0) {
      for (f in sub_files) {
        subchrplot_filename <- gsub("ChrPlot_", "", tools::file_path_sans_ext(basename(f)))
        subchrplot_filename <- gsub("subCX_", "", subchrplot_filename)
        subchrplot_filename <- gsub("_", " ", subchrplot_filename)
        cat("##### ", subchrplot_filename, "\n\n", sep = "")
        emit_grid(f, cols = 1, max_height_px = params$plot_max_height_px)
      }
    }
  }
}

if (isTRUE(params$run_TEs_distance_n_size)) {
  cat("#### TE methylation vs TE size & centromere distance\n\n")
  emit_dir(TEs_distance_n_size_path, recursive = TRUE, cols = 4, max_height_px = params$plot_max_height_px)
}

if (isTRUE(params$total_meth_annotation)) {
  cat("#### Total methylation across genome annotations {.tabset}\n\n")
  files <- list.files(total_meth_annotation_path, pattern = "\\.csv$", full.names = TRUE, ignore.case = TRUE)
  if (length(files) == 0) {
    cat("_No total methylation annotation tables found._\n\n")
  } else {
    for (csv in files) {
      cat("##### ", tools::file_path_sans_ext(basename(csv)), "\n\n", sep = "")
      tab <- tryCatch(read.csv(csv, nrows = 50, check.names = FALSE), error = function(e) NULL)
      if (!is.null(tab)) {
        kable(tab)
        cat("\n\n")
      } else {
        cat("_Could not read table._\n\n")
      }
    }
  }
}

if (isTRUE(params$run_TF_motifs)) {
  cat("#### TF motif analysis\n\n")
  emit_dir(TF_motifs_path, recursive = TRUE, cols = params$default_plot_cols, max_height_px = params$plot_max_height_px)
}
})}
```

```{r metaplot-tes, results='asis'}
if (isTRUE(params$run_TE_metaPlots) || isTRUE(params$run_GeneBody_metaPlots) || isTRUE(params$run_GeneFeatures_metaPlots)) { safe_section({
  cat("### MetaPlots {.tabset}\n\n")

  if (isTRUE(params$run_TE_metaPlots)) {
    cat("#### TEs\n\n")
    emit_dir(file.path(meta_dir, "TEs"), recursive = FALSE, cols = 4, max_height_px = params$plot_max_height_px)
  }

  if (isTRUE(params$run_GeneBody_metaPlots)) {
    cat("#### Genes\n\n")
    emit_dir(file.path(meta_dir, "Genes"), recursive = FALSE, cols = 4, max_height_px = params$plot_max_height_px)
  }

  if (isTRUE(params$run_GeneFeatures_metaPlots)) {
    cat("#### Gene features\n\n")
    emit_dir(file.path(meta_dir, "Gene_features"), recursive = FALSE, cols = 3, max_height_px = params$plot_max_height_px)
  }
})}
```

```{r dmrs-section, results='asis'}
if (isTRUE(params$analyze_DMRs)) { safe_section({
  cat("### DMRs {.tabset}\n\n")
  cat("#### Settings\n\n")

settings <- data.frame(
  Setting = c("methyl_files_type","binSize (bp)","minCytosinesCount","minReadsPerCytosine","pValueThreshold","minProportionDiff (CG,CHG,CHH)"),
  Value = c(val1(params$methyl_files_type), val1(params$binSize), val1(params$minCytosinesCount), val1(params$minReadsPerCytosine),
            val1(params$pValueThreshold), val1(params$minProportionDiff)),
  Description = c(
    "Type of input methylation files (e.g., CX_report).",
    "Genome tiling window size used to reduce per-site noise.",
    "Minimum informative cytosines required per bin.",
    "Per-site coverage filter applied before summarizing bins.",
    "Statistical cutoff for significance (p-value).",
    "Minimal ΔmC thresholds for CG, CHG and CHH contexts."
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)
knitr::kable(settings)

    ## Density
    dens <- list.files(DMRs_analysis_path, pattern = "DMRs_Density.*\\.(png|svg|jpg|jpeg|tiff)$", full.names = TRUE, ignore.case = TRUE)
    if (length(dens) > 0) {
        cat("#### Density\n\n")
        emit_grid(dens, cols = 1, max_height_px = params$plot_max_height_px)
    }

    ## Gain/Loss
    cat("#### Gain/Loss\n\n")
    gl_files <- list_media_files(gainORloss_path, recursive = TRUE)
    pie_files <- gl_files[grepl("pie", basename(gl_files), ignore.case = TRUE)]
    other_files <- setdiff(gl_files, pie_files)

    if (length(other_files) > 0) {
        emit_grid(other_files, cols = 3, max_height_px = params$plot_max_height_px / 1.85)
    }
    if (length(pie_files) > 0) {
        emit_grid(pie_files, cols = 3, max_height_px = params$plot_max_height_px / 6)
    }

    ## Chromosome distribution
    cat("#### Chromosome plots\n\n")
    emit_dir(ChrPlots_DMRs_path, recursive = TRUE, cols = 1, max_height_px = params$plot_max_height_px / 3)

    ## Genome annotation {.tabset}
    cat("#### Genome annotation {.tabset}\n\n")

    ## Main genome annotation plots
    ## Show two main genome annotation images as separate tabs
    ann_files <- list.files(genome_ann_path, full.names = TRUE)
    ann_files <- ann_files[file.exists(ann_files) & !file.info(ann_files)$isdir]
    ann_files <- ann_files[sapply(ann_files, is_media)]
    ## Identify the two main annotation files by pattern
    cg_file <- ann_files[grepl("DMRs_genome_annotation", basename(ann_files), ignore.case = TRUE)]
    te_vs_cg_file <- ann_files[grepl("DMRs_TEs_vs_coding_genes", basename(ann_files), ignore.case = TRUE)]
    rest_files <- setdiff(ann_files, c(cg_file, te_vs_cg_file))
    if (length(cg_file) > 0) {
        cat("##### Coding genes\n\n")
        emit_grid(cg_file, cols = 1, max_height_px = params$plot_max_height_px)
    }
    if (length(te_vs_cg_file) > 0) {
        cat("##### TEs vs coding genes\n\n")
        emit_grid(te_vs_cg_file, cols = 1, max_height_px = params$plot_max_height_px / 1.8)
    }
    if (length(rest_files) > 0) {
        for (f in order_files(rest_files)) {
            cat("##### ", tools::file_path_sans_ext(basename(f)), "\n\n", sep = "")
            emit_grid(c(f), cols = 1, max_height_px = params$plot_max_height_px)
        }
    }

    ## TEs additional results tab
    te_add_dir <- file.path(genome_ann_path, "TEs_addiotionnal_results")
    if (dir.exists(te_add_dir)) {
        cat("##### TEs additional results\n\n")
        ## List and display the two main images
        te_imgs <- c("TEs_lines_density_superfamilies", "TEs_direction_density_superfamilies")
        te_imgs <- unlist(lapply(te_imgs, function(base) {
            list_media_files(te_add_dir, pattern = paste0("^", base, "\\.(png|svg|jpg|jpeg|tiff)$"), recursive = FALSE)
        }))
        if (length(te_imgs) > 0) {
            emit_grid(te_imgs, cols = 2, max_height_px = params$plot_max_height_px)
        }

        ## Display the three super-family frequency plots in one tab, three columns
        sff_dir <- file.path(te_add_dir, "super_family_frequency")
        sff_imgs <- list.files(sff_dir, pattern = "DMRs_.*\\.(png|svg|jpg|jpeg|tiff)$", full.names = TRUE, ignore.case = TRUE)
        if (length(dens) > 0) {
            cat("##### TE super-family frequency\n\n")
            emit_grid(sff_imgs, cols = 3, max_height_px = params$plot_max_height_px)
        }
    }

    ## Functional groups tabset
    if (isTRUE(params$run_functional_groups)) {
        fg_dir <- file.path(genome_ann_path, "functional_groups")
        if (dir.exists(fg_dir)) {
            cat("##### Functional groups {.tabset}\n\n")
            fg_files <- list_media_files(fg_dir, recursive = TRUE)
            if (length(fg_files) == 0) {
                cat("_No functional group plots found._\n\n")
            } else {
                ## Remove *_groups_volcano_ pattern
                fg_files <- fg_files[!grepl("_groups_volcano_", basename(fg_files), ignore.case = TRUE)]
                ## Move *_groups_barPlots_ to front and rename to 'summary'
                bar_idx <- grepl("_groups_barPlots_", basename(fg_files), ignore.case = TRUE)
                bar_files <- fg_files[bar_idx]
                rest_files <- fg_files[!bar_idx]
                ## For bar files, replace pattern with ' summary'
                bar_files_named <- setNames(
                    bar_files,
                    gsub(paste0("_groups_barPlots_", params$var2, "_vs_", params$var1), " summary", tools::file_path_sans_ext(basename(bar_files)), ignore.case = TRUE)
                )
                ## Try to group by context keywords in filenames
                fg_cg <- rest_files[grepl("\\bcg\\b", basename(rest_files), ignore.case = TRUE)]
                fg_chg <- rest_files[grepl("\\bchg\\b", basename(rest_files), ignore.case = TRUE)]
                fg_chh <- rest_files[grepl("\\bchh\\b", basename(rest_files), ignore.case = TRUE)]
                fg_rest <- setdiff(rest_files, c(fg_cg, fg_chg, fg_chh))
                ## Show summary tab first if present
                if (length(bar_files_named) > 0) {
                    for (i in seq_along(bar_files_named)) {
                        cat("#####", names(bar_files_named)[i], "\n\n")
                        emit_grid(bar_files_named[i], cols = params$default_plot_cols, max_height_px = params$plot_max_height_px)
                    }
                }
                if (length(fg_cg) > 0) {
                    cat("###### CG\n\n")
                    emit_grid(fg_cg, cols = params$default_plot_cols, max_height_px = params$plot_max_height_px)
                }
                if (length(fg_chg) > 0) {
                    cat("###### CHG\n\n")
                    emit_grid(fg_chg, cols = params$default_plot_cols, max_height_px = params$plot_max_height_px)
                }
                if (length(fg_chh) > 0) {
                    cat("###### CHH\n\n")
                    emit_grid(fg_chh, cols = params$default_plot_cols, max_height_px = params$plot_max_height_px)
                }
                if (length(fg_rest) > 0) {
                    for (f in order_files(fg_rest)) {
                        cat("###### ", tools::file_path_sans_ext(basename(f)), "\n\n", sep = "")
                        emit_grid(c(f), cols = 1, max_height_px = params$plot_max_height_px)
                    }
                }
            }
        }
    }

    ## BigWigs
    if (dir.exists(DMRs_bigWig_path)) {
        bw <- list.files(DMRs_bigWig_path, pattern = "\\.(bw|bigWig)$", full.names = FALSE, ignore.case = TRUE)
        if (length(bw) > 0) {
            cat("#### BigWig tracks\n\n")
            cat("These tracks can be loaded in IGV.\n\n")
            cat(paste0("- ", bw, collapse = "\n"), "\n\n")
        }
    }

    ## GO / KEGG (plots only)
    if (isTRUE(params$run_GO_analysis)) {
        cat("#### GO analysis\n\n")
        emit_dir(file.path(exp_path, "GO_analysis"), recursive = TRUE, cols = params$default_plot_cols, max_height_px = params$plot_max_height_px)
    }
    if (isTRUE(params$run_KEGG_pathways)) {
        cat("#### KEGG pathways\n\n")
        emit_dir(file.path(exp_path, "KEGG_pathway"), recursive = TRUE, cols = params$default_plot_cols, max_height_px = params$plot_max_height_px)
    }
})}
```

```{r strand-asymmetry, results='asis'}
## Helper function to capitalize first letter
capitalize <- function(x) { paste0(toupper(substring(x, 1, 1)), substring(x, 2)) }

if (isTRUE(params$analyze_strand_asymmetry_DMRs)) {
  safe_section({
  cat("### Strand-asymmetry DMRs {.tabset}\n\n")

  for (cntx in c("CG", "CHG", "CHH")) {
    cat("#### ", cntx, " {.tabset}\n\n", sep = "")
    for (pat in c("conflicting", "hemi", "symmetric")) {
      files <- list.files(file.path(strand_asymmetry_path, cntx), pattern = paste0("DMRs_", cntx, "_", pat, ".*\\.csv$"), full.names = TRUE, ignore.case = TRUE)
      for (file in files) {
        if (file.exists(file)) {
          cat(sprintf("##### %s\n\n", capitalize(pat)))
          tab <- read.csv(file, nrows = 50, check.names = FALSE)
          kable(tab)
          cat("\n\n")
        }
      }
    }
  }
  })
}
```

```{r other, results='asis'}
if (isTRUE(params$analyze_DMVs) || isTRUE(params$analyze_dH)) {
  cat("### Other analyses\n\n")
}

if (isTRUE(params$analyze_DMVs)) { safe_section({
  cat("#### DMVs (1kb) {.tabset}\n\n")
  cat("Below are the first 50 rows of each DMV table (if present). Full tables are available in the output directory.\n\n")
  if (dir.exists(DMV_analysis_path)) {
    csvs <- list.files(DMV_analysis_path, pattern = "\\.csv$", full.names = TRUE, ignore.case = TRUE)
    if (length(csvs) > 0) {
      for (csv in csvs) {
        cat("##### ", tools::file_path_sans_ext(basename(csv)), "\n\n", sep = "")
        tab <- tryCatch(read.csv(csv, nrows = 50, check.names = FALSE), error = function(e) NULL)
        if (!is.null(tab)) {
          kable(tab)
          cat("\n\n")
        } else {
          cat("_Could not read table._\n\n")
        }
      }
    } else {
      cat("_No DMV tables found._\n\n")
    }
    emit_dir(DMV_analysis_path, recursive = TRUE, cols = params$default_plot_cols)
  }
})}

if (isTRUE(params$analyze_dH)) { safe_section({
  cat("#### ΔH / entropy\n\n")
  cat("ΔH captures changes in methylation heterogeneity (stochasticity) rather than only mean methylation.\n\n")
  emit_dir(dH_path, recursive = TRUE, cols = params$default_plot_cols)
  if (dir.exists(dH_ann_path)) {
    cat("##### ΔH genome annotation\n\n")
    emit_dir(dH_ann_path, recursive = TRUE, cols = params$default_plot_cols)
  }
})}
```

### Session {.tabset}

#### Configurations {.tabset}

```{r params-table, results='asis'}
pnames3 <- names(params)[1:5]
pnames2 <- names(params)[8:13]
pnames1 <- names(params)[c(6:7, 14:length(params))]

vals3 <- vapply(pnames3, function(n) val1(params[[n]]), character(1))
vals2 <- vapply(pnames2, function(n) val1(params[[n]]), character(1))
vals1 <- vapply(pnames1, function(n) val1(params[[n]]), character(1))

decrip3 <- c(
  "Maximum plot width (pixels)",
  "Maximum plot height (pixels)",
  "Gap between plots in grids (pixels)",
  "Default columns for multi-plot sections",
  "If TRUE: copy plots next to the HTML report"
)

decrip2 <- c(
  "Path to first sample methylation file",
  "Path to second sample methylation file",
  "Path to Methylome.Plants project root",
  "Genome Annotation file",
  "Description file",
  "Transposable Elements file"
)

decrip1 <- c(
  "First sample name",
  "Second sample name",
  "Minimum proportion difference for CG, CHG, CHH",
  "DMRs bin size",
  "Minimum cytosines count",
  "Minimum reads per cytosine",
  "P-value (padj) threshold",
  "Post-alignment file type ('CX_report', 'bedMethyl', 'CGmap')",
  "Output images format",
  "Number of cores",
  "Perform the main DMRs analysis workflow",
  "Sample-level QC plots (distribution, scatter, correlation, variance)",
  "Perform PCA for total methylation levels",
  "Total methylation bar-plot",
  "Total methylation chromosome plot",
  "Analyze TEs total methylation by size and distance from centromere",
  "Total methylation per genic annotations",
  "Transcription factors motif analysis",
  "Functional groups genes overlap DMRs",
  "Perform GO analysis over DMRs",
  "Perform KEGG pathways analysis over DMRs",
  "Analyze strand-specific (+/-) DMRs",
  "Analyze differentially methylated valleys (1kbp)",
  "Analyze SurpDMRs (delta-H = -(p * log2(p) + (1 - p) * log2(1 - p)))",
  "Analyze TEs metaPlots",
  "Analyze Genes-body metaPlots",
  "Analyze Gene Features metaPlots",
  "Bin-size for 'Gene_features' metaPlots",
  "Number of random genes/TEs for metaPlots ('all' for all coding-genes and TEs)"
)
cfg3 <- data.frame(Parameter = pnames3, Value = vals3, Description = decrip3, check.names = FALSE, stringsAsFactors = FALSE)
cfg2 <- data.frame(Parameter = pnames2, Value = vals2, Description = decrip2, check.names = FALSE, stringsAsFactors = FALSE)
cfg1 <- data.frame(Parameter = pnames1, Value = vals1, Description = decrip1, check.names = FALSE, stringsAsFactors = FALSE)

rownames(cfg3) <- NULL
rownames(cfg2) <- NULL
rownames(cfg1) <- NULL

cat("##### Pipeline parameters\n\n")
kable(cfg1)
cat("##### Data and annotation files\n\n")
kable(cfg2)
cat("##### Report parameters\n\n")
kable(cfg3)
```

#### .log file

```{r session-log, results='asis'}
log_path <- list.files(file.path(exp_path), pattern = paste0(".*\\.log$"), full.names = TRUE, ignore.case = TRUE)[1]
if (!is.na(log_path) && file.exists(log_path)) {
  log_lines <- readLines(log_path, warn = FALSE)
  colorize_log <- function(lines) {
    # Use Unicode private-use placeholders so regex never touches generated HTML
    O  <- "\U0091"   # stands for <
    C  <- "\U0092"   # stands for >
    wrap <- function(txt, css) paste0(O, "span style='", css, "'", C, txt, O, "/span", C)

    out <- vapply(lines, function(line) {
      # 1. Escape real HTML entities
      line <- gsub("&", "&amp;", line, fixed = TRUE)
      line <- gsub("<", "&lt;",  line, fixed = TRUE)
      line <- gsub(">", "&gt;",  line, fixed = TRUE)

      # 2. Section dividers (---- ... ----) → dark bold
      if (grepl("^-{4,}", line)) {
        return(wrap(line, "color:#555; font-weight:bold;"))
      }

      # 3. Timestamps inside [] → blue bold
      line <- gsub("\\[([0-9:]+)\\]",
                    paste0("[", wrap("\\1", "color:#007acc; font-weight:bold;"), "]"),
                    line)

      # 4. "Error" / "Warning" keywords
      line <- gsub("\\bError\\b",   wrap("Error",   "color:#cb0a0a; font-weight:bold;"), line)
      line <- gsub("\\bWarning\\b", wrap("Warning", "color:#e6a700; font-weight:bold;"), line)

      # 5. Percentages and plain numbers (not inside entities/tags)
      line <- gsub("(?<![&#;a-zA-Z])\\b([0-9]+\\.?[0-9]*%)\\b",
                    wrap("\\1", "color:#0b7d3e; font-weight:bold;"), line, perl = TRUE)
      line <- gsub("(?<![&#;a-zA-Z])\\b([0-9]+\\.?[0-9]*)\\b",
                    wrap("\\1", "color:#d32f2f;"), line, perl = TRUE)

      # 6. "done" → green
      line <- gsub("\\bdone\\b", wrap("done", "color:#0b7d3e;"), line)

      line
    }, character(1))

    # Convert placeholders to real angle brackets at the very end
    result <- paste(out, collapse = "\n")
    result <- gsub(O, "<", result, fixed = TRUE)
    result <- gsub(C, ">", result, fixed = TRUE)
    result
  }
  cat("<div style='max-width:900px; margin:0 auto; background:#f9f9f9; border:1px solid #ddd; border-radius:6px; box-shadow:0 2px 8px #eee;'><pre style='font-size:90%; color:#222; padding:16px; line-height:1.5; overflow-x:auto;'>",
      colorize_log(log_lines),
      "</pre></div>", sep = "")
} else {
  cat("_No .log file found._\n\n")
}
```

#### Session info

```{r session-info, results='asis'}
si <- capture.output(sessionInfo())
colorize_si <- function(lines) {
  O <- "\U0091"; C <- "\U0092"
  wrap <- function(txt, css) paste0(O, "span style='", css, "'", C, txt, O, "/span", C)
  out <- vapply(lines, function(line) {
    line <- gsub("&", "&amp;", line, fixed = TRUE)
    line <- gsub("<", "&lt;",  line, fixed = TRUE)
    line <- gsub(">", "&gt;",  line, fixed = TRUE)
    # Section headers (e.g. "R version ...", "attached base packages:", etc.)
    if (grepl("^(R version|Platform|Running under|locale|time zone|attached base packages|other attached packages|loaded via a namespace)", line)) {
      return(wrap(line, "color:#007acc; font-weight:bold;"))
    }
    # Package names with version: name_version
    line <- gsub("\\b([a-zA-Z][a-zA-Z0-9.]*_[0-9][0-9.\\-]*)\\b", wrap("\\1", "color:#555;"), line)
    line
  }, character(1))
  result <- paste(out, collapse = "\n")
  result <- gsub(O, "<", result, fixed = TRUE)
  result <- gsub(C, ">", result, fixed = TRUE)
  result
}
cat("<div style='max-width:900px; margin:0 auto; background:#f9f9f9; border:1px solid #ddd; border-radius:6px; box-shadow:0 2px 8px #eee;'><pre style='font-size:90%; color:#222; padding:16px; line-height:1.5; overflow-x:auto;'>",
    colorize_si(si),
    "</pre></div>", sep = "")
```
